---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uisapi

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/uisapi)](https://cran.r-project.org/package=uisapi)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/uisapi)](https://cran.r-project.org/package=uisapi)
![R CMD Check](https://github.com/tidy-intelligence/r-uisapi/actions/workflows/R-CMD-check.yaml/badge.svg)
![Lint](https://github.com/tidy-intelligence/r-uisapi/actions/workflows/lint.yaml/badge.svg)
[![Codecov test coverage](https://codecov.io/gh/tidy-intelligence/r-uisapi/graph/badge.svg)](https://app.codecov.io/gh/tidy-intelligence/r-uisapi)
<!-- badges: end -->

Retrieve data from the [UNESCO Institute for Statistics (UIS) API](https://api.uis.unesco.org/api/public/documentation/). UIS provides public access to more than 4,000 indicators focusing on education, science and technology, culture, and communication.

The package is part of the [econdataverse](https://www.econdataverse.org/) family of packages aimed at helping economists and financial professionals work with sovereign-level economic data.

 > 💡 The UIS API has no known rate limiting, but there is a 100,000 record limit for each request. For larger data downloads, the UIS recommends its [Bulk Data Download Service (BDDS)](https://databrowser.uis.unesco.org/resources/bulk).

## Installation

You can install `uisapi` from [CRAN](https://cran.r-project.org/package=uisapi) via:

```{r eval=FALSE}
install.packages("uisapi")
```

You can install the development version of `uisapi` from GitHub with:

```{r eval=FALSE}
# install.packages("pak")
pak::pak("tidy-intelligence/r-uisapi")
```

## Usage

Load the package:

```{r}
library(uisapi)
```

To fetch indicators for specific entities, you can call:

```{r}
uis_get(
  entities = c("BRA", "USA"),
  indicators = c("CR.1", "CR.2"),
  start_year = 2010,
  end_year = 2020
)
```

You can get information about the available entities via:

```{r}
uis_get_entities()
```

The list of available indicators can be retrieved via:

```{r}
uis_get_indicators()
```

The API supports versioning and you can retrieve all versions using:

```{r}
uis_get_versions()
```

If you are only interested in the current default version, you can use the parameter `default_only`:

```{r}
uis_get_versions(default_only = TRUE)
```

The API will only return 100,000 records for each query, if more data is requested, the API call with fail with a 400 http status code. If you need more data, then UIS recommends using the [Bulk Data Download Service (BDDS)](https://databrowser.uis.unesco.org/resources/bulk). You can get a list of available files via:

```{r}
uis_bulk_files()
```

Each bulk files consists of multiple CSV files and the `uisapi` package currently does not provide automatic importers for these files. Rather, you can use base utils to download a specific file, unzip it, and read its files. For instance: 

```{r eval = FALSE}
download.file(
  "https://uis.unesco.org/sites/default/files/documents/bdds/092024/SDG.zip",
  destfile = "SDG.zip"
)
unzip("SDG.zip")
read.csv("SDG_COUNTRY.csv")
```

## Contributing

Contributions to `uisapi` are welcome! If you'd like to contribute, please follow these steps:

1. **Create an issue**: Before making changes, create an issue describing the bug or feature you're addressing.
2. **Fork the repository**: After receiving supportive feedback from the package authors, fork the repository to your GitHub account.
3. **Create a branch**: Create a branch for your changes with a descriptive name.
4. **Make your changes**: Implement your bug fix or feature.
5. **Test your changes**: Run tests to ensure your changes don't break existing functionality.
6. **Submit a pull request**: Push your changes to your fork and submit a pull request to the main repository.
